Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-04-25 paper/ms.tex
Formatting updates after Lars edits

  Elongated bar-like features are ubiquitous in galaxies, occurring at the centers of
  approximately two-thirds of spiral disks \cite{2000AJ....119..536E,
  2007ApJ...657..790M}.  Due to gravitational interactions between the bar and
  the other components of galaxies, it is expected that angular momentum and
  matter will redistribute over long (Gyr)
  timescales in barred galaxies \cite{1972MNRAS.157....1L,
  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas phase of galaxies has conclusively 
  demonstrated that bars should slow their rotation over time 
  due to their interaction with dark matter halos
  \cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
  2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
  2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
  2009ApJ...697..293D}.  We have performed a simulation of a Milky Way-like
  galactic disk hosting a strong bar which includes a state-of-the-art model
  of the interstellar medium and a live dark matter halo. In this simulation
  the bar pattern does not slow down over time, and instead remains at a
  stable, constant rate of rotation. This behavior has been observed in
  previous simulations using more simplified models for the 
  interstellar gas but its explanation has remained
  elusive\cite{1993AA...268...65F, 2010ApJ...719.1470V}. We propose that the
  gas phase of the disk and the dark matter halo act in concert to stabilize
  the bar pattern speed and prevent a bar from slowing down or speeding up. This result
  naturally explains why nearly all observed bars rotate
  rapidly\cite{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G}
  and is especially relevant for our understanding of how the Milky Way
  arrived at its present state.

>>>>

  Elongated bar-like features are ubiquitous in galaxies, occurring at the
  centers of approximately two-thirds of spiral
  disks\cite{2000AJ....119..536E, 2007ApJ...657..790M}.  Due to gravitational
  interactions between the bar and the other components of galaxies, it is
  expected that angular momentum and matter will redistribute over long (Gyr)
  timescales in barred galaxies \cite{1972MNRAS.157....1L,
  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas
  phase of galaxies has conclusively demonstrated that bars should slow their
  rotation over time due to their interaction with dark matter halos
  \cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
  2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
  2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
  2009ApJ...697..293D}. We have performed a simulation of a Milky Way-like
  galactic disk hosting a strong bar which includes a state-of-the-art model
  of the interstellar medium and a live dark matter halo. In this simulation
  the bar pattern does not slow down over time, and instead remains at a
  stable, constant rate of rotation. This behavior has been observed in
  previous simulations using more simplified models for the interstellar gas
  but its explanation has remained elusive\cite{1993AA...268...65F,
  2010ApJ...719.1470V}. We propose that the gas phase of the disk and the dark
  matter halo act in concert to stabilize the bar pattern speed and prevent a
  bar from slowing down or speeding up. This result naturally explains why
  nearly all observed bars rotate rapidly\cite{2011MSAIS..18...23C,
  2015AA...576A.102A, 2019MNRAS.482.1733G} and is especially relevant for our
  understanding of how the Milky Way arrived at its present state.
